---
title: "Process and Clean Climate Data"
author: "Michelle Evans"
date: ""
output: html_document
---

This document processes data logger data used in the `urbanDengue` project. 

The clean data is output into `data/microclimate/clean/2016TrialsAdult.csv` (and as RData).

```{r setup}
library(dplyr)
```

# Total Climate

## Merge CSVs 

This works from inside the microclimate data folder, reading files from `raw` and saving in `clean`.

```{r load these grouped tables}
aug1 <- read.csv("denguePaper/data/microclimate/raw/AdultTable_July14-Aug8_2016.csv")[,-1]
aug2 <- read.csv("denguePaper/data/microclimate/raw/AdultTable_Aug8-Aug29_2016.csv")[,-1]
aug3 <- read.csv("denguePaper/data/microclimate/raw/AdultTable_Aug29-Sep19_2016.csv")[,-1]
oct1 <- read.csv("denguePaper/data/microclimate/raw/AdultTable_Sep19-Oct14_2016.csv")
oct2 <- read.csv("denguePaper/data/microclimate/raw/AdultTable_Oct14-Nov7_2016.csv")
loggerAll <- rbind(aug1, aug2, aug3, oct1, oct2)
#correct Date format
loggerAll$Date <- as.character(loggerAll$Date)
loggerAll$Date <- strptime(loggerAll$Date, format="%m/%d/%Y %H:%M:%S")

#cutoff those before august start date
climate <- loggerAll[loggerAll$Date>strptime("08-01-2016 17:00:00", format="%m-%d-%Y %H:%M:%S"),]

```

## Clean and Format

```{r}
#toss out ridiculous levels
climate <- climate[climate$Temp<65,]
#format date
climate$Date <- strptime(climate$Date, format="%Y-%m-%d %H:%M:%S")
#get date
climate$Day <- as.Date(climate$Date)
#drop horrible format of date
climate <- select(climate, -Date)

# add tray id to climate data
trayID <- read.csv("denguePaper/data/microclimate/trayLoggerID.csv") #read in IDs
climate <- merge(climate, trayID, by="Pot_ID")

#fix duplicates because some loggers were downloaded multiple times
climate <- unique(climate)
```

## Drop malfunctioning loggers

U2T2, U2T4, R3T4

```{r}
toDrop <- which(climate$Tray_ID %in% c("U2T2", "U2T4", "R3T4"))
climate <- climate[-toDrop,]
rm(toDrop)
```

## Save

```{r}
saveRDS(climate, "denguePaper/data/microclimate/clean/2016TrialsAdult.RData")
write.csv(climate, "denguePaper/data/microclimate/clean/2016TrialsAdult.csv", row.names = F)
```

# Replicate Summary Statistics for Climate Stats

```{r}
climate <- readRDS("denguePaper/data/microclimate/clean/2016TrialsAdult.RData")
```

Calculate the necessary statistics per the summer and fall feplicate to be used in the climate statistical analysis.

This calculates the mean daily mean, minimum, maximum and range per tray over both blocks.

```{r}
fall <- climate %>%
  filter(Day>"2016-09-26") %>%
  mutate(Block="Fall")

summer <- climate %>%
  filter(Day < "2016-09-01") %>%
  mutate(Block="Summer")

climBlocks <- rbind(summer, fall)
rm(summer,fall)

climTray <- climBlocks %>%
  dplyr::group_by(Tray_ID, Day, Site_ID, Class, Block) %>%
  dplyr::summarise(meanT=mean(Temp, na.rm = T), 
                   minT=min(Temp, na.rm = T), 
                   maxT=max(Temp, na.rm = T), 
                   meanRH=mean(RH, na.rm = T), 
                   maxRH=max(RH, na.rm = T),
                   minRH=min(RH, na.rm = T), 
                   DTR = max(Temp, na.rm = T) - min(Temp, na.rm = T),
                   DHR = max(RH, na.rm = T) - min(RH, na.rm = T)) %>% #values per tray
  ungroup() %>%
  group_by(Block, Class, Site_ID, Tray_ID) %>% #get mean values per class
  dplyr::summarise(meanT=mean(meanT), minT=mean(minT), maxT=mean(maxT), meanRH=mean(meanRH), maxRH=mean(maxRH), minRH=mean(minRH), meanDTR = mean(DTR), meanDHR = mean(DHR)) %>%
  ungroup()

climTray$Block <- factor(climTray$Block, levels = c("Summer", "Fall"))
```

Impute mean for missing data loggers:

```{r}
allLoggers <- data.frame(expand.grid(
  Tray_ID = levels(climate$Tray_ID),
  Block = levels(climTray$Block)), 
  Site_ID = rep(rep(levels(climate$Site_ID), each =4),2), 
  Class = rep(rep(levels(climate$Class), each = 12), 2))

climImpute <- right_join(climTray, allLoggers, by = c("Block", "Tray_ID", "Class", "Site_ID")) 

climImpute <- climImpute %>%
  group_by(Block, Class, Site_ID) %>%
  mutate_all(funs(impute(., mean))) %>%
  ungroup()

climTray <- climImpute
```


```{r}
saveRDS(climTray, "denguePaper/data/microclimate/clean/climateTraySummary.RData")
```

# Weighted Climate Data

This calculates climate data that is weighted by the number of mosquitoes that emerged from each tray that were used for infections. It is a mean calculated on the individual emerging mosquitoes, rather than each tray. For example, on site U3, no mosquitoes emerged from one tray becuase it was knocked over by wildlife, therefore its microclimate data isn't included in the weighted mean used to calculate the effects of climate on infection.

This data is only to be used in reference to infections.

## Summer

```{r}
augEmerg <- read.csv("denguePaper/data/emergence/raw/AugustEmergence.csv")

augInf <- augEmerg %>%
  filter(Sex=="F") %>%
    filter((Site_Code %in% c("U2", "U1", "S3", "R1") & Day <= 14) |
           (Site_Code %in% c("S1", "S2", "R2", "R3", "U3") & Day <= 17)) %>%
  mutate(Block="Summer")
```

```{r}
augClim <- climate %>%
  filter(Day >= as.Date("2016-08-01","%Y-%m-%d")) %>%
    filter((Site_ID %in% c("U2", "U1", "S3", "R1") & Day <= "2016-08-14") |
           (Site_ID %in% c("S1", "S2", "R2", "R3", "U3") & Day <= "2016-08-17")) %>%
  dplyr::select(-Site_ID, -Pot_ID, -Class) %>%
  #get daily averages by Tray
  group_by(Tray_ID, Day) %>%
  summarise_all(funs(mean(., na.rm=T), min(., na.rm=T), max(., na.rm=T))) %>%
  #calculate DTR
  mutate(DTR=Temp_max-Temp_min) %>%
  #get overall average over study period per tray (average daily values)
  ungroup() %>%
  dplyr::select(-Day) %>%
  group_by(Tray_ID) %>%
  summarise_all(funs(mean))

##new method of weighting climate by expanded infection
augInfExp <- augInf[rep(seq.int(1,nrow(augInf)), augInf$Num_Emerge),]
test <- merge(augInfExp, augClim, by.x="Tray_Code", by.y="Tray_ID", all.x=T)

augEnvVar <- test %>%
  dplyr::select(-Tray_Code,-Site, -Tray, - Class, - Month, - Day, - Exp_Day, - Sex, -Num_Emerge) %>%
  group_by(Block, Site_Code) %>%
  summarise_all(funs(mean(.,na.rm=T),se=(sd(., na.rm=T)/sqrt(n())))) %>%
  ungroup() %>%
  rename(Site_ID = Site_Code)

rm(test)
```

## Fall

```{r}
octEmerg <- read.csv("denguePaper/data/emergence/raw/OctoberEmergence.csv")

octInf <- octEmerg %>%
  filter(Sex=="F") %>%
  filter((Site_Code %in% c("U3", "R3") & Day <= 21) |
           (Site_Code %in% c("U2", "U1", "S3") & Day <= 20) |
           (Site_Code %in% c("S1", "S2", "R1", "R2") & Day <= 24)) %>%
  mutate(Block = "Fall")
```

```{r}
octClim <- climate %>%
  filter(Day >= as.Date("2016-09-26","%Y-%m-%d")) %>%
  #filter out appropriate days
  filter((Site_ID %in% c("U3", "R3") & Day <= "2016-10-21") |
           (Site_ID %in% c("U2", "U1", "S3") & Day <= "2016-10-20") |
           (Site_ID %in% c("S1", "S2", "R1", "R2") & Day <= "2016-10-24")) %>%
  dplyr::select(-Site_ID, -Pot_ID, -Class) %>%
  #get daily averages by Tray
  group_by(Tray_ID, Day) %>%
  summarise_all(funs(mean(., na.rm=T), min(., na.rm=T), max(., na.rm=T))) %>%
  #calculate DTR
  mutate(DTR=Temp_max-Temp_min) %>%
  #get overall average over study period per tray (average daily values)
  ungroup() %>%
  dplyr::select(-Day) %>%
  group_by(Tray_ID) %>%
  summarise_all(funs(mean))

##new method of weighting climate by expanded infection
octInfExp <- octInf[rep(seq.int(1,nrow(octInf)), octInf$Num_Emerge),]
test <- merge(octInfExp, octClim, by.x="Tray_Code", by.y="Tray_ID", all.x=T)

octEnvVar <- test %>%
  dplyr::select(-Tray_Code,-Site, -Tray, - Class, - Month, - Day, - Exp_Day, - Sex, -Num_Emerge) %>%
  group_by(Block,Site_Code) %>%
  summarise_all(funs(mean(.,na.rm=T),se=(sd(., na.rm=T)/sqrt(n())))) %>%
  ungroup() %>%
  rename(Site_ID = Site_Code)
  
rm(test) #clear unused temporary dataframe
```

## Save

```{r}
saveRDS(augEnvVar, "denguePaper/data/microclimate/clean/summerInfectionClimate.RData")
saveRDS(octEnvVar, "denguePaper/data/microclimate/clean/fallInfectionClimate.RData")
```

