---
title: "EvansCode"
author: "Michelle Evans"
date: ""
output: html_document
---

#Set-Up

```{r knitr setup}
knitr::opts_chunk$set(echo = T)
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(warning=FALSE)
```

```{r package and color setup}
library(xtable)
library(tidyr)
library(ggplot2) 
library(lme4) #mixed models
library(car)
library(MASS)
library(gridExtra) #for facet grids later
library(cowplot)
library(caret)
library(ggthemes)
library(survival)
library(DHARMa) #mixed model residuals
library(survminer) #extracts survival objects
library(dplyr)

colR <- "dodgerblue"
colS <- "gray80"
colU <- "maroon"
```

# Data Loading and Formatting

Load and format infection data:
```{r data loading and formatting}
formatData <- function(month){
  #' format infection data
  #' @params month (ie. "august")
  #' @returns dataframe of properly formatted data
  #adjust wingLength
  monthDf <- read.csv(paste0("../data/infections/raw/", month,"Dengue.csv"))
  #convert wingLength and drop extra columns
  monthDf$Wing <- monthDf$WingLength*monthDf$conversion..mm.bar.
  monthDf <- dplyr::select(monthDf, -WingLength, -conversion..mm.bar.)
  
  #dpi as factor
  monthDf$DPI <- as.factor(monthDf$DPI)
  
  #add in class and site
  monthDf$site <- as.factor(substr(as.character(monthDf$Individual), 1, 2))
  monthDf$class <- NULL
  for (i in 1:nrow(monthDf)){
    if (substr(monthDf$site[i], 1,1)=="R"){
    monthDf$class[i] <- "Rural"
    } else if (substr(monthDf$site[i], 1,1)=="S"){
    monthDf$class[i] <- "Suburban"
    } else if (substr(monthDf$site[i], 1,1)=="U"){
    monthDf$class[i] <- "Urban"
    }
  }
  monthDf$class <- as.factor(monthDf$class)
  
  #convert Y and N to 1 and 0 for statistics
  levels(monthDf$Body) <- c("NA", 0, 1)
  monthDf$Body <- as.numeric(as.character(monthDf$Body))
  levels(monthDf$Saliva) <- c("NA", 0, 1)
  monthDf$Saliva <- as.numeric(as.character(monthDf$Saliva))
  # august had no contaminated heads, so different corrections
  if (month=="august"){
    levels(monthDf$Head) <- c(0, 1)
  } else levels(monthDf$Head) <- c("NA",0, 1)
  monthDf$Head <- as.numeric(as.character(monthDf$Head))
  
  ##Fix false negatievs
  #adjust so that if saliva is positive, so is head
  #ddjust so that is head is positive, so is body
  monthDf$Head[monthDf$Saliva>0] <- 1
  monthDf$Body[monthDf$Head>0] <- 1
  
  return(monthDf)
}

august <- formatData("august")
oct <- formatData("october")
seasons <- rbind(august,oct)
seasons$block <- as.factor(c(rep("summer", nrow(august)), rep("fall", nrow(oct))))

seasonSumm <- seasons %>%
  filter(DPI==21) %>%
  #drop individual
  dplyr::select(-Individual, -site, -Wing, -DPI) %>%
  group_by(block, class) %>%
  summarise_each(funs(mean(.,na.rm=T),sd(.,na.rm=T),se=(sd(., na.rm=T)/sqrt(n())))) %>%
  ungroup()
```

Load and format microclimate data
```{r load and format microclimate data}
climate <- read.csv('../data/microclimate/clean/2016TrialsAdult.csv')[,-1]
#toss out ridiculous levels
climate <- climate[climate$Temp<75,]
#format date
climate$Date <- strptime(climate$Date, format="%Y-%m-%d %H:%M:%S")
#draw out day
climate$Day <- as.Date(climate$Date)

# add tray id to climate data
trayID <- read.csv("../data/microclimate/trayLoggerID.csv") #read in IDs
climate <- merge(climate, trayID, by="Pot_ID")

#fix duplicates for R1T1
climate <- unique(climate)

#U2T2 and U1T2 are missing data
# range(climate[climate$Tray_ID=="U2T2", 'Date']) 
# range(climate[climate$Tray_ID=="U1T2", 'Date']) 

#drop U2T2 becuase it only has data until August 5th
inds <- which(climate$Tray_ID=="U2T2")
climate <- climate[-inds,]
rm(inds)

#U2T4 wasn't working right, reporting temps above 40C in October
inds <- which(climate$Tray_ID=="U2T4")
climate <- climate[-inds,]
rm(inds)
```

Load and format emergence data
```{r load emergence data,echo=FALSE}
augEmerg <- read.csv("../data/emergence/raw/AugustEmergence.csv")
augEmerg$block <- as.factor("summer")
octEmerg <- read.csv("../data/emergence/raw/OctoberEmergence.csv")
octEmerg$block <- as.factor("fall")

allEmerg <- rbind(augEmerg, octEmerg)
```

Crop and standardize emergence data to be used with infections
```{r standardize and correct for emergence, echo=FALSE}
#filter out only days when infected mosquitoes were in the trays
octInf <- octEmerg %>%
  filter(Sex=="F") %>%
  filter((Site_Code %in% c("U3", "R3") & Day <= 21) |
           (Site_Code %in% c("U2", "U1", "S3") & Day <= 20) |
           (Site_Code %in% c("S1", "S2", "R1", "R2") & Day <= 24))

augInf <- augEmerg %>%
  filter(Sex=="F") %>%
    filter((Site_Code %in% c("U2", "U1", "S3", "R1") & Day <= 14) |
           (Site_Code %in% c("S1", "S2", "R2", "R3", "U3") & Day <= 17))
```

```{r process october climate, message=F, echo=FALSE}
octClim <- climate %>%
  dplyr::select(-Date) %>%
  filter(Day >= as.Date("2016-09-26","%Y-%m-%d")) %>%
  #filter out appropriate days
  filter((Site_ID %in% c("U3", "R3") & Day <= "2016-10-21") |
           (Site_ID %in% c("U2", "U1", "S3") & Day <= "2016-10-20") |
           (Site_ID %in% c("S1", "S2", "R1", "R2") & Day <= "2016-10-24")) %>%
  dplyr::select(-Site_ID, -Pot_ID, -Class) %>%
  #get daily averages by Tray
  group_by(Tray_ID, Day) %>%
  summarise_each(funs(mean(., na.rm=T), min(., na.rm=T), max(., na.rm=T))) %>%
  #calculate DTR
  mutate(DTR=Temp_max-Temp_min) %>%
  #get overall average over study period per tray (average daily values)
  ungroup() %>%
  dplyr::select(-Day) %>%
  group_by(Tray_ID) %>%
  summarise_each(funs(mean))

##new method of weighting climate by expanded infection
octInfExp <- octInf[rep(seq.int(1,nrow(octInf)), octInf$Num_Emerge),]
test <- merge(octInfExp, octClim, by.x="Tray_Code", by.y="Tray_ID", all.x=T)

octEnvVar <- test %>%
  dplyr::select(-Tray_Code,-Site, -Tray, - Class, - Month, - Day, - Exp_Day, - Sex, -Num_Emerge) %>%
  group_by(block,Site_Code) %>%
  summarise_each(funs(mean(.,na.rm=T),se=(sd(., na.rm=T)/sqrt(n())))) 

  
rm(test) #clear unused temporary dataframe
```

```{r august climate formatting, message=F, echo=FALSE}
##repeat above for August
augClim <- climate %>%
  dplyr::select(-Date) %>%
  filter(Day >= as.Date("2016-08-01","%Y-%m-%d")) %>%
    filter((Site_ID %in% c("U2", "U1", "S3", "R1") & Day <= "2016-08-14") |
           (Site_ID %in% c("S1", "S2", "R2", "R3", "U3") & Day <= "2016-08-17")) %>%
  dplyr::select(-Site_ID, -Pot_ID, -Class) %>%
  #get daily averages by Tray
  group_by(Tray_ID, Day) %>%
  summarise_each(funs(mean(., na.rm=T), min(., na.rm=T), max(., na.rm=T))) %>%
  #calculate DTR
  mutate(DTR=Temp_max-Temp_min) %>%
  #get overall average over study period per tray (average daily values)
  ungroup() %>%
  dplyr::select(-Day) %>%
  group_by(Tray_ID) %>%
  summarise_each(funs(mean))

##new method of weighting climate by expanded infection
augInfExp <- augInf[rep(seq.int(1,nrow(augInf)), augInf$Num_Emerge),]
test <- merge(augInfExp, augClim, by.x="Tray_Code", by.y="Tray_ID", all.x=T)

augEnvVar <- test %>%
  dplyr::select(-Tray_Code,-Site, -Tray, - Class, - Month, - Day, - Exp_Day, - Sex, -Num_Emerge) %>%
  group_by(block,Site_Code) %>%
  summarise_each(funs(mean(.,na.rm=T),se=(sd(., na.rm=T)/sqrt(n())))) 

rm(test)

```

```{r, echo=FALSE}
#merge the summer and fall into one dataframe
oct$block <- as.factor("fall")
august$block <- as.factor("summer")
seasonInf <- merge(rbind(oct, august[august$DPI=="21",]), rbind(augEnvVar,octEnvVar), by.x=c("block", "site"), by.y=c("block", "Site_Code"))
```

# Infection Dynamics

## Infection by Class and Season

Function to Plot Infections
```{r season plot function}
seasonPlot <- function(bodyPart, summTable=seasonSumm){
  cols <- c(1,2,grep(bodyPart, colnames(summTable)))
  tempLarge <- summTable[,cols]
  colnames(tempLarge)[3:5] <- c("mean", "sd", "se")
    
  tempMean <- tempLarge %>%
    dplyr::select(-sd,-se) %>%
    spread(block, mean) %>%
    dplyr::select(-class)
  
  #put summer before fall
  tempMean <- tempMean[,c(2,1)]

  
  rownames(tempMean) <- levels(tempLarge$class)
  tempMean <- t(as.matrix(tempMean))
  
  colvec <- c(rep(colR,2), rep(colS,2), 
                            rep(colU,2))
  
  barCenters <- barplot(tempMean, 
                      beside=T,
                      col=colvec,
                      density = c(40,NA),
                      names.arg=colnames(tempMean),
                      ylim=c(0,0.80)
                      )
  
  tempSE <- tempLarge %>%
    dplyr::select(-sd,-mean) %>%
    spread(block, se) %>%
    dplyr::select(-class)
  
   #put summer before fall
  tempSE <- tempSE[,c(2,1)]
  
   tempSE <- t(tempSE)
  
  #add se bars
  segments(barCenters, tempMean - tempSE , barCenters,
         tempMean + tempSE, lwd = 1.5)
  
  arrows(barCenters, tempMean - tempSE , barCenters,
         tempMean + tempSE,
         lwd = 1.5, 
         angle = 90,
         code = 3, length = 0.03)
  
  #legend("topright", legend=c("Summer", "Fall"), density = c(40,NA),
       #bty = "n", col = "gray80")
}
```

Bar graph of infection by season and land class:
```{r season by class infection plot, fig.width=4, fig.height=6}
#png(file="../figures/forMS/seasonInfections.png", width = 4, height=6, units="in", res=500, family="sans")
par(mfrow=c(3,1), las=1, par(mar=c(5.1,6.1,4.1,2.1)))
seasonPlot(bodyPart="Body")
par(xpd=T)
legend(x=7.5, y=1.0, legend=c("Summer", "Fall"), density = c(40,NA),
       bty = "n", col = "gray80", cex=1.1)
mtext("Body")



seasonPlot(bodyPart="Head")
mtext("Head")
par(las=0)
mtext("Proportion Infected", side=2, line=3)



par(las=1)
seasonPlot(bodyPart="Saliva")
mtext("Saliva")
#dev.off()
```

Statistics on infection by season and land class
```{r seasonal infection stats}
#model #there was no interaction of class and block
mixModelseasonsBody21 <- lme4::glmer(Body~class + block + (1|site), 
                          data=seasons[seasons$DPI=="21",],
                          family=binomial(link="logit"))
Body21Aov <- car::Anova(mixModelseasonsBody21) #Wald test
#confint(mixModelseasonsBody21) #profiled confidence interval
drop1(mixModelseasonsBody21, test="Chisq") #Likelihood ratio test

mixModelseasonsHead21 <- lme4::glmer(Head~class + block  + (1|site), 
                          data=seasons[seasons$DPI=="21",],
                          family=binomial(link="logit"))
Head21Aov <- car::Anova(mixModelseasonsHead21) #Wald test
#confint(mixModelseasonsHead21) #profiled confidence interval
drop1(mixModelseasonsHead21, test="Chisq")

mixModelseasonsSaliva21 <- lme4::glmer(Saliva~class + block  + (1|site), 
                          data=seasons[seasons$DPI=="21",],
                          family=binomial(link="logit"))
summary(mixModelseasonsSaliva21) #z-test
Saliva21Aov <- car::Anova((mixModelseasonsSaliva21)) #Wald test
#confint(mixModelseasonsSaliva21) #profiled confidence interval
drop1(mixModelseasonsSaliva21, test="Chisq")
```

### Infection Efficiency

Format data for infection efficiency

```{r create and format efficiency data}
bodyEff <- seasons %>%
  filter(DPI==21) %>%
  #drop individual
  select(-Individual, -site, -Wing, -DPI) %>%
  group_by(block, class) %>%
  summarise_each(funs(bodyMean=mean(.,na.rm=T),bodySE=(sd(., na.rm=T)/sqrt(n()))), Body) %>%
  ungroup()

headEff <- seasons %>%
  filter(DPI==21) %>%
  filter(Body==1) %>%
  #drop individual
  select(-Individual, -site, -Wing, -DPI) %>%
  group_by(block, class) %>%
  summarise_each(funs(headMean=mean(.,na.rm=T),headSE=(sd(., na.rm=T)/sqrt(n()))), Head) %>%
  ungroup()

salEff <-  seasons %>%
  filter(DPI==21) %>%
  filter(Head==1) %>%
  #drop individual
  select(-Individual, -site, -Wing, -DPI) %>%
  group_by(block, class) %>%
  summarise_each(funs(salMean=mean(.,na.rm=T),salSE=(sd(., na.rm=T)/sqrt(n()))), Saliva) %>%
  ungroup()

#group together
allEff <- full_join(bodyEff, headEff, by=c("block", "class"))
allEff <- full_join(allEff, salEff, by=c("block", "class"))

library(tidyr)
meltMean <- allEff %>%
  select(block, class, contains("Mean")) %>%
  gather(key=variable, value=mean, -block, -class) 

meltMean$type <- rep(c("Body", "Head", "Saliva"), each=6)
  
meltSE <- allEff %>%
  select(block, class, contains("SE")) %>%
  gather(variable, SE, -block, -class)

meltSE$type <- rep(c("Body", "Head", "Saliva"), each=6)
  
meltAll <- full_join(meltMean, meltSE, by=c("class", "block", "type")) %>%
  select(-variable.x, -variable.y)
```

Plot of infection efficiency
```{r plot efficiency}
#supplemental plot
ggplot(data=meltAll[order(meltAll$block, decreasing=F),], aes(x=class, group=block))+
  geom_bar(stat="identity", aes(y=mean, alpha=factor(block), fill=factor(class)), color="gray20", position=position_dodge(width=-0.9))+
  facet_wrap(~type, nrow=3, dir="v") +
  scale_fill_manual(values=c(colR, colS, colU))+
  scale_alpha_discrete(range=c(1,0.3), name="Season", labels=c("Fall", "Summer"), guide=guide_legend(reverse=T))+
  geom_errorbar(aes(ymin=mean-SE, ymax=mean+SE), width=0.2, color="gray20", position =position_dodge(-0.9))+
  theme_base() + 
  xlab("Land Class") +
  ylab("Infection Efficiency") +
  guides(fill=F)
  #theme(legend.title=element_text("Season"))
```

Infection efficiency statistics
```{r efficiency stats}
#model 
mixModelseasonsBody21 <- lme4::glmer(Body~class * block + (1|site), 
                          data=seasons[seasons$DPI=="21",],
                          family=binomial(link="logit"))
Body21Aov <- car::Anova(mixModelseasonsBody21)

mixModelseasonsHead21 <- lme4::glmer(Head~class * block  + (1|site), 
                          data=seasons[seasons$DPI=="21" & seasons$Body==1,],
                          family=binomial(link="logit"))
Head21Aov <- car::Anova(mixModelseasonsHead21) #no effect

mixModelseasonsSaliva21 <- lme4::glmer(Saliva~class * block  + (1|site), 
                          data=seasons[seasons$DPI=="21" & seasons$Head==1,],
                          family=binomial(link="logit"))
Saliva21Aov <- car::Anova((mixModelseasonsSaliva21))
```

## Infection and Microclimate

```{r create dataframe of site-aggregated infections}
seasonSite <- seasons %>%
  filter(DPI=="21") %>%
  dplyr::select(-Individual, -DPI, - Wing) %>%
  dplyr::group_by(block, class, site) %>%
  summarise_each(funs(mean(.,na.rm=T), se=(sd(., na.rm=T)/sqrt(n()))))

#group with temperature data
seasonInfSite <- merge(seasonSite, rbind(augEnvVar,octEnvVar), by.x=c("block", "site"), by.y=c("block", "Site_Code"))

# inspect variable correlations
# cor(seasonInfSite[,10:18])
```

```{r body x climate stats}
respV <- "Body_mean"
predVs <- c("Temp_mean_mean","RH_mean_mean")

myCols <- c(respV, predVs, "block","class", "site")
modDF <- seasonInfSite %>%
  dplyr::select(one_of(myCols)) 
modDF <- na.omit(modDF)

#standardize IVs
modDF$Temp_mean_mean <- scale(modDF$Temp_mean_mean, center=T, scale=T)
modDF$RH_mean_mean <- scale(modDF$RH_mean_mean, center=T, scale=T)

finalBody <- lm(Body_mean ~ Temp_mean_mean+RH_mean_mean, data=modDF)
qqPlot(finalBody, labels=modDF$site)
plot(cooks.distance(finalBody))
abline(h=4/length(modDF), lty=2)
car::Anova(finalBody)
#confint(finalBody)
summary(finalBody)

outlierTest(finalBody)
leveragePlots(finalBody)
vif(finalBody)
sqrt(vif(finalBody))>2
```

```{r head x climate stats}
respV <- "Head_mean"
predVs <- c("Temp_mean_mean","RH_mean_mean")

myCols <- c(respV, predVs, "block","class", "site")
modDF <- seasonInfSite %>%
  dplyr::select(one_of(myCols)) 
modDF <- na.omit(modDF)

#standardize IVs
modDF$Temp_mean_mean <- scale(modDF$Temp_mean_mean, center=T, scale=T)
modDF$RH_mean_mean <- scale(modDF$RH_mean_mean, center=T, scale=T)

finalHead <- lm(Head_mean ~ Temp_mean_mean+RH_mean_mean, data=modDF)
qqPlot(finalHead, labels=modDF$site)
plot(cooks.distance(finalHead))
abline(h=4/length(modDF), lty=2)
car::Anova(finalHead)
#confint(finalHead)
summary(finalHead)

outlierTest(finalHead)
leveragePlots(finalHead)
```

```{r body x climate stats}
respV <- "Saliva_mean"
predVs <- c("Temp_mean_mean","RH_mean_mean")

myCols <- c(respV, predVs, "block","class", "site")
modDF <- seasonInfSite %>%
  dplyr::select(one_of(myCols)) 
modDF <- na.omit(modDF)

#standardize IVs
modDF$Temp_mean_mean <- scale(modDF$Temp_mean_mean, center=T, scale=T)
modDF$RH_mean_mean <- scale(modDF$RH_mean_mean, center=T, scale=T)

finalSal <- lm(Saliva_mean ~ Temp_mean_mean+RH_mean_mean, data=modDF)
qqPlot(finalSal, labels=modDF$site)
plot(cooks.distance(finalSal))
abline(h=4/length(modDF), lty=2)
car::Anova(finalSal)
#confint(finalSal)
summary(finalSal)

outlierTest(finalSal)
leveragePlots(finalSal)
```

Infection x Temperature Plot

```{r three panel temp x infection plot}
#png(file="figures/forMS/InfxTemp3Panel.png", width = 7, height=3, units="in", res=500, family="sans")
plotBody <- ggplot(data=seasonInfSite, aes(x=Temp_mean_mean, y=Body_mean))+
  geom_errorbar(aes(ymin=Body_mean-Body_se, ymax=Body_mean+Body_se), color="gray20", width=0.1) +
  #geom_errorbarh(aes(xmin=Temp_mean_mean-Temp_mean_se, xmax=Temp_mean_mean+Temp_mean_se, y=Body_mean),color="gray20", height=0.05) + #not enough error to see
  geom_point(aes(color=class, shape=block), size=4.5) + 
  scale_color_manual(values=c(colR, colS, colU)) +
  theme_fivethirtyeight() +
    theme(panel.background = element_rect(fill = "transparent", colour = NA), 
        plot.background = element_rect(fill = "transparent", colour = NA),
        legend.key = element_blank())+
  theme(legend.background = element_rect(fill = "transparent", colour = NA)) +
    theme(axis.title = element_text(), axis.title.x = element_text(), axis.title.y=element_blank()) +
  ylab("Prop. Infected")+
  xlab("") +  
  theme(legend.position="none")

plotHead <- ggplot(data=seasonInfSite, aes(x=Temp_mean_mean, y=Head_mean))+
   geom_errorbar(aes(ymin=Head_mean-Head_se, ymax=Head_mean+Head_se), color="gray20", width=0.1) +
  #geom_errorbarh(aes(xmin=Temp_mean_mean-Temp_mean_se, xmax=Temp_mean_mean+Temp_mean_se), color="gray20", height=0.05) +
  geom_point(aes(color=class, shape=block), size=4) + 
  scale_color_manual(values=c(colR, colS, colU)) +
  theme_fivethirtyeight() +
    theme(panel.background = element_rect(fill = "transparent", colour = NA), 
        plot.background = element_rect(fill = "transparent", colour = NA),
        legend.key = element_blank())+
  theme(legend.background = element_rect(fill = "transparent", colour = NA)) +
    theme(axis.title = element_text(), axis.title.x = element_text(), axis.title.y=element_blank())+
  #ylab("Prop. Disseminated")+
  xlab("Mean Daily Temperature (C)")+
  theme(legend.position="none")

plotSaliva <- ggplot(data=seasonInfSite, aes(x=Temp_mean_mean, y=Saliva_mean))+
  geom_errorbar(aes(ymin=Saliva_mean-Saliva_se, ymax=Saliva_mean+Saliva_se), color="gray20", width=0.1) +
  #geom_errorbarh(aes(xmin=Temp_mean_mean-Temp_mean_se, xmax=Temp_mean_mean+Temp_mean_se, y=Body_mean),color="gray20", height=0.05) + #not enough error to see
  geom_point(aes(color=class, shape=block), size=4.5) + 
  scale_color_manual(values=c(colR, colS, colU)) +
  theme_fivethirtyeight() +
    theme(panel.background = element_rect(fill = "transparent", colour = NA), 
        plot.background = element_rect(fill = "transparent", colour = NA),
        legend.key = element_blank())+
  theme(legend.background = element_rect(fill = "transparent", colour = NA)) +
    theme(axis.title = element_text(), axis.title.x = element_text(), axis.title.y=element_blank()) +
  #ylab("Prop. Infectious")+
  xlab("") +  
  scale_y_continuous(limits=c(0,1))+
  #theme(legend.position="right", legend.direction="vertical")
  theme(legend.position="none")


plot_grid(plotBody, plotHead, plotSaliva, labels=c("A", "B", "C"), ncol=3)
#dev.off()
```

Infection x Relative Humidity Plot
```{r}
#png(file="figures/forMS/InfxRH3Panel.png", width = 7, height=3, units="in", res=500, family="sans")
plotBody <- ggplot(data=seasonInfSite, aes(x=RH_mean_mean, y=Body_mean))+
  geom_errorbar(aes(ymin=Body_mean-Body_se, ymax=Body_mean+Body_se), color="gray20", width=0.1) +
  #geom_errorbarh(aes(xmin=RH_mean_mean-RH_mean_se, xmax=RH_mean_mean+RH_mean_se, y=Body_mean),color="gray20", height=0.05) + #not enough error to see
  geom_point(aes(color=class, shape=block), size=4.5) + 
  scale_color_manual(values=c(colR, colS, colU)) +
  theme_fivethirtyeight() +
    theme(panel.background = element_rect(fill = "transparent", colour = NA), 
        plot.background = element_rect(fill = "transparent", colour = NA),
        legend.key = element_blank())+
  theme(legend.background = element_rect(fill = "transparent", colour = NA)) +
    theme(axis.title = element_text(), axis.title.x = element_text(), axis.title.y=element_blank()) +
  ylab("Prop. Infected")+
  xlab("") +  
  theme(legend.position="none")

plotHead <- ggplot(data=seasonInfSite, aes(x=RH_mean_mean, y=Head_mean))+
   geom_errorbar(aes(ymin=Head_mean-Head_se, ymax=Head_mean+Head_se), color="gray20", width=0.1) +
  #geom_errorbarh(aes(xmin=RH_mean_mean-RH_mean_se, xmax=RH_mean_mean+RH_mean_se), color="gray20", height=0.05) +
  geom_point(aes(color=class, shape=block), size=4) + 
  scale_color_manual(values=c(colR, colS, colU)) +
  theme_fivethirtyeight() +
    theme(panel.background = element_rect(fill = "transparent", colour = NA), 
        plot.background = element_rect(fill = "transparent", colour = NA),
        legend.key = element_blank())+
  theme(legend.background = element_rect(fill = "transparent", colour = NA)) +
    theme(axis.title = element_text(), axis.title.x = element_text(), axis.title.y=element_blank())+
  #ylab("Prop. Disseminated")+
  xlab("Mean Relative Humidity (%)")+
  theme(legend.position="none")

plotSaliva <- ggplot(data=seasonInfSite, aes(x=RH_mean_mean, y=Saliva_mean))+
  geom_errorbar(aes(ymin=Saliva_mean-Saliva_se, ymax=Saliva_mean+Saliva_se), color="gray20", width=0.1) +
  #geom_errorbarh(aes(xmin=RH_mean_mean-RH_mean_se, xmax=RH_mean_mean+RH_mean_se, y=Body_mean),color="gray20", height=0.05) + #not enough error to see
  geom_point(aes(color=class, shape=block), size=4.5) + 
  scale_color_manual(values=c(colR, colS, colU)) +
  theme_fivethirtyeight() +
    theme(panel.background = element_rect(fill = "transparent", colour = NA), 
        plot.background = element_rect(fill = "transparent", colour = NA),
        legend.key = element_blank())+
  theme(legend.background = element_rect(fill = "transparent", colour = NA)) +
    theme(axis.title = element_text(), axis.title.x = element_text(), axis.title.y=element_blank()) +
  #ylab("Prop. Infectious")+
  xlab("") +  
  scale_y_continuous(limits=c(0,1))+
  #theme(legend.position="right", legend.direction="vertical")
  theme(legend.position="none")


plot_grid(plotBody, plotHead, plotSaliva, labels=c("A", "B", "C"), ncol=3)
#dev.off()
```

## Infection and Body Size

### Body Infection

```{r body inf x size stats}
bodyWingDF <- seasonInf %>%
  dplyr::select(Body, Wing, site, block, class) %>%
  filter(!is.na(Body)) %>%
  filter(!is.na(Wing))


bodyWing <- glmer(Body~Wing + (1|site),
                          data=bodyWingDF,
                          family=binomial(link="logit"))
summary(bodyWing)
require(DHARMa)
simulationOutput <- simulateResiduals(fittedModel = bodyWing, n=99)
plotSimulatedResiduals(simulationOutput = simulationOutput)
#plotResiduals(bodyWingDF$Wing,  simulationOutput$scaledResiduals)
#confint(bodyWing)
Anova(bodyWing)
```

```{r body inf x size plot}
ggplot(data=bodyWingDF, aes(x=factor(Body), y=Wing))+
  geom_boxplot() +
  geom_jitter(shape=16, position=position_jitter(0.1), aes(color=factor(class)), alpha=0.5) +
    theme_fivethirtyeight()+
  theme(axis.title = element_text()) +
  theme(panel.background = element_rect(fill = "transparent", colour = NA), 
        plot.background = element_rect(fill = "transparent", colour = NA),
        legend.key = element_blank())+
  theme(legend.background = element_rect(fill = "transparent", colour = NA))+
  theme(legend.position="right", legend.direction="vertical") +
  scale_color_manual(values=c("dodgerblue", "gray10", "maroon"), labels=c("Rural", "Suburban", "Urban"))+
    scale_fill_manual(values=c("dodgerblue", "gray80", "maroon"), labels=c("Rural", "Suburban", "Urban")) +
  guides(color=F) +
  xlab("Body Positivity") +
  ylab("Wing Size")
```

### Head Infection

```{r head inf x size stats}
headWingDF <- seasonInf %>%
  dplyr::select(Head, Wing, site, block, class) %>%
  filter(!is.na(Head)) %>%
  filter(!is.na(Wing))


headWing <- glmer(Head~Wing + (1|site),
                          data=headWingDF,
                          family=binomial(link="logit"))
summary(headWing)
simulationOutput <- simulateResiduals(fittedModel = headWing, n=99)
plotSimulatedResiduals(simulationOutput = simulationOutput)
#plotResiduals(headWingDF$Wing,  simulationOutput$scaledResiduals)
Anova(headWing)
```

### Saliva Infection

```{r saliva inf x size stats}
salWingDF <- seasonInf %>%
  dplyr::select(Saliva, Wing, site, block, class) %>%
  filter(!is.na(Saliva)) %>%
  filter(!is.na(Wing))


salWing <- glmer(Saliva~Wing + (1|site),
                          data=salWingDF,
                          family=binomial(link="logit"))
summary(salWing)
simulationOutput <- simulateResiduals(fittedModel = salWing, n=99)
plotSimulatedResiduals(simulationOutput = simulationOutput)
#plotResiduals(salWingDF$Wing,  simulationOutput$scaledResiduals)
```

# Emergence and Survival

## Data Format and Loading

```{r}
emergAug <- read.csv("../data/emergence/raw/AugustEmergence.csv")
emergAug$block <- "summer"

emergOct <- read.csv("../data/emergence/raw/OctoberEmergence.csv")
emergOct$block <- "fall"
#drop U3T1 in fall because it was eaten by ants
emergOct <- filter(emergOct, Tray_Code!="U3T1")
emergOct <- filter(emergOct, Tray_Code!="S1T3") #dumped
emergOct <- filter(emergOct, Tray_Code!="U1T4") #dumped

emergAll <- rbind(emergOct, emergAug)
#expand so each mosquito gets one row
emergExp <- emergAll[rep(seq.int(1,nrow(emergAll)), emergAll$Num_Emerge),
                      c(11, 2,4,5,6:9)]
#sum(emergAll$Num_Emerge)==nrow(emergExp) #quick check this worked

#get survival per tray
survSumm <- emergExp %>%
  filter(Sex=="F") %>%
  group_by(block, Tray_Code) %>%
  dplyr::mutate(percSurv=n()) %>%
  ungroup() %>%
  dplyr::select(block, Class, Site_Code, Tray_Code, percSurv)
survSumm <- unique(survSumm)
```

```{r fillIn function}
fillIn <- function(df, endDay, totalMosq=50){
  #' Fill In Emergence Dates
  #' this function fills in for those mosquitoes that did not emerge so we do not have data for, it gives them an observation/event of 0 on the last day we found a mosquito emerged
  #' @param df the data frame you wish to fill in, in our case by pot
  #' @param endDay the last day of emergence
  #' @param totalMosq estimated starting number of mosquitoes per pot
  #' @returns dataframe with census data filled in for mosquitoes that did not emerge
  toRep <- df[1,]
  toRep$Exp_Day <- endDay
  toRep$event <- 0
  if(nrow(df)<totalMosq){
    toAdd <- toRep[rep(1, (totalMosq-nrow(df))),]
    allTest <- rbind(df, toAdd)
  } else {
    toAdd <- NA
    allTest <- NA
  }
  return(allTest)
}
```

```{r apply FillIn function}
applyFill <- function(season, allData=emergExp){
  #' Apply FillIn function
  #' @param season "fall" or "summer"
  #' @param allData full dataframe with row for each mosquito that emerged
  #' @returns censused data for the full season
  tempList <- list()
  tempDF <- allData
  tempDF <- tempDF[tempDF$Sex=="F",]
  tempDF <- tempDF[tempDF$block==season,]
  tempDF$event <- 1 #add emergence event
  for (i in 1:length(unique(tempDF$Tray_Code))){ 
    df <- tempDF[tempDF$Tray_Code==unique(tempDF$Tray_Code)[i],]
    endDay <- max(tempDF$Exp_Day)
    tempList[[i]] <- fillIn(df=df, endDay=endDay)
  }
  allSurv <- do.call(rbind.data.frame, tempList)
  return(allSurv)
}
```

```{r}
summerSurv <- applyFill(season="summer")
fallSurv <- applyFill(season="fall")
allSurv <- rbind(summerSurv, fallSurv)
```

## Class x Block

### Survival

```{r survival stats}
siteSumm <- emergExp %>%
  filter(Sex=="F") %>%
  group_by(block, Tray_Code) %>%
  dplyr::mutate(numSurv=n()) %>%
  ungroup() %>%
  dplyr::select(block, Class, Site_Code, numSurv)
siteSumm <- unique(siteSumm)

siteSumm$block <- as.factor(siteSumm$block)

siteEmerg <- lmer(numSurv ~ 0 + Class + block + (1|Site_Code), data=siteSumm)
summary(siteEmerg)
plot(siteEmerg)

#pairwise
summary(multcomp::glht(siteEmerg, linfct = multcomp::mcp(block = "Tukey"), test = multcomp::adjusted("holm")))

summary(multcomp::glht(siteEmerg, linfct = multcomp::mcp(Class = "Tukey"), test = multcomp::adjusted("holm")))
```

### Emergence

```{r}
emergExpF <- filter(emergExp, Sex=="F")

emergExpF$block <- as.factor(emergExpF$block)

mixEmerge <- lmer(Exp_Day ~  0 + Class + block + (1|Site_Code), data=emergExpF)
plot(mixEmerge)
qqnorm(resid(mixEmerge))
summary(mixEmerge)
#confint(mixEmerge)

#pairwise
summary(glht(mixEmerge, linfct = mcp(block = "Tukey"), test = adjusted("holm")))

summary(glht(mixEmerge, linfct = mcp(Class = "Tukey"), test = adjusted("holm")))
```

### Plots

Adjust factors for plotting:
```{r}
levels(emergExpF$block) <- c("Fall", "Summer")
emergExpF <- mutate(emergExpF, block=fct_rev(block))
levels(emergExpF$Class) <- c("Rural", "Suburban", "Urban")

levels(siteSumm$block) <- c("Fall", "Summer")
siteSumm <- mutate(siteSumm, block=fct_rev(block))
levels(siteSumm$Class) <- c("Rural", "Suburban", "Urban")
```


Survival and Emergence Plot

```{r}
#png(file="figures/forMS/emergSurvPlots.png", width = 4, height=6, units="in", res=500, family="sans")
#Survival plot
survPlot <- ggplot(data=siteSumm, aes(x=Class, y=(numSurv*2)))+
  geom_boxplot(aes(fill=Class, alpha=0.5), width=0.4)+
  scale_fill_manual(values=c(colR, colS, colU), labels=c("Rural", "Suburban", "Urban"))+
  facet_wrap(~block, ncol=2)+
  ylab("Percent Survial")+
  theme(legend.position="none")+
  theme(axis.title = element_text(size=12))+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())

#Emergence Plot
emergePlot <- ggplot(emergExpF, aes(x=Class, y=(1/Exp_Day)))+
  geom_boxplot(aes(fill=Class, alpha=0.5), width=0.4)+
  scale_fill_manual(values=c(colR, colS, colU), labels=c("Rural", "Suburban", "Urban"))+
  facet_wrap(~block, ncol=2)+
  ylab("Development\nRate")+
  theme(legend.position="none") +
  theme(strip.background = element_blank(),
       strip.text.x = element_blank()) +
  xlab("Land Class") +
  theme(axis.text.x = element_text(size=8)) +
  theme(axis.title = element_text(size=12), axis.title.y= element_text(size=12))

#grid together
plot_grid(survPlot, emergePlot, labels=c("A", "B"), nrow=2)

#dev.off()
```


## Microclimate

Tie in tray-level climate data:
```{r}
allClim <- rbind(augClim, octClim)
allClim$block <- rep(c("summer", "fall"), each=34)
survClim <- merge(survSumm, allClim, by.x=c("block","Tray_Code"), by.y=c("block","Tray_ID"))
```

### Survival

Statistics
```{r}
modDF <- survClim
#standardize environmental variables
modDF$Temp_mean <- scale(modDF$Temp_mean, center=T, scale=T)
modDF$RH_mean <- scale(modDF$RH_mean, center=T, scale=T)
survClimModmix <- lmer(percSurv~Temp_mean+(1|Site_Code), #RH has no significant effect
                    data=survClim)

plot(survClimModmix)
qqnorm(resid(survClimModmix))
summary(survClimModmix)
coef(survClimModmix)
AIC(survClimModmix)

anova(survClimModmix)
```


### Development Rate

```{r}
allClim$block <- as.factor(allClim$block)
levels(allClim$block) <- c("Fall", "Summer")
emergClim <- merge(emergExpF, allClim, by.x=c("block","Tray_Code"), by.y=c("block","Tray_ID"))
emergClim <- mutate(emergClim, devRate =1/Exp_Day)
```

Statistics:
```{r}
survClimMod <- lmer(devRate~Temp_mean*RH_mean +(1|Site_Code),
                    data=emergClim)

plot(survClimMod)
qqnorm(resid(survClimMod))
summary(survClimMod)
coef(survClimMod)

car::Anova(survClimMod)
```

#Body Size

Load body size data:
```{r load body size data}
augWing <- read.csv("../data/emergence/raw/AugustWingLength.csv", stringsAsFactors = F)
octWing <- read.csv("../data/emergence/raw/OctoberWingLength.csv",  stringsAsFactors = F)


#convert to mm & clean
augWing$mm <- augWing$Bars*augWing$Conversion.mm.bars.
octWing$mm <- octWing$Bars*octWing$Conversion.bars.mm.

octWing$site <- as.factor(substr(as.character(octWing$TrayCode), 1, 2))
augWing$site <- as.factor(substr(as.character(augWing$TrayCode), 1, 2))

getClass <- function(monthDf){
  monthDf$class <- NULL
  for (i in 1:nrow(monthDf)){
    if (substr(monthDf$site[i], 1,1)=="R"){
    monthDf$class[i] <- "Rural"
    }
    if (substr(monthDf$site[i], 1,1)=="S"){
    monthDf$class[i] <- "Suburban"
    }
    if (substr(monthDf$site[i], 1,1)=="U"){
    monthDf$class[i] <- "Urban"
    }
  }
  monthDf$class <- as.factor(monthDf$class)
  return(monthDf)
}

augWing <- getClass(augWing)
octWing <- getClass(octWing)

octWing$block <- "fall"
augWing$block <- "summer"

augWing$Date <- as.Date(as.character(augWing$Date), format="%m/%d/%Y")
octWing$Date <- as.Date(as.character(octWing$Date), format="%m/%d/%Y")

#add day of experiment
augWing$Exp_Day <- as.numeric(augWing$Date-as.Date("2016-08-01", format="%Y-%m-%d"))
octWing$Exp_Day <- as.numeric(octWing$Date-as.Date("2016-09-26", format="%Y-%m-%d"))

#rename traycode column to match
colnames(augWing)[1] <- "Tray_ID"
colnames(octWing)[1] <- "Tray_ID"

#combine
fallWing <- octWing %>%
  dplyr::select(block, class, site, Tray_ID, Exp_Day, mm)

summerWing <- augWing %>%
  dplyr::select(block, class, site, Tray_ID, Exp_Day, mm)

allWing <- rbind(summerWing, fallWing)
allWing$block<- as.factor(allWing$block)

#drop outlier in S2 (wing size =1.56 mm)
allWing <- allWing %>%
  filter(mm>1.6)
```

Statistics:
```{r}
wingMix <- lmer(mm~0 + class*block +(1|site),
                data=allWing)
summary(wingMix)
plot(wingMix)
car::Anova(wingMix)
coef(wingMix)
```

```{r}
allWingSumm <- summarySE(allWing, measurevar="mm", groupvars=c("class", "block"))

ggplot(data=allWingSumm, aes(x=class, y=mm, group=block))+
  geom_point(aes(y=mm, shape=factor(block), color=factor(class)), position=position_dodge(width=-0.9), size=5)+
  scale_color_manual(values=c(colRural, colSuburban, colU))+
  scale_alpha_discrete(range=c(1,0.3), name="Season", labels=c("Fall", "Summer"), guide=guide_legend(reverse=T))+
  scale_shape_discrete(name="Season",labels=c("Fall", "Summer"), guide=guide_legend(reverse=T))+
  geom_errorbar(aes(ymin=mm-se, ymax=mm+se), width=0.2, color="gray20", position =position_dodge(-0.9))+
  ylim(2.2,2.6)+
  theme_fivethirtyeight() +
  theme(panel.background = element_rect(fill = "transparent", colour = NA), 
        plot.background = element_rect(fill = "transparent", colour = NA),
        legend.key = element_blank())+
  theme(legend.background = element_rect(fill = "transparent", colour = NA)) +
    theme(axis.title = element_text(), axis.title.x = element_text()) +
  ylab("Uninfected Wing Size")+
  xlab("Land Class") +  
  #ggtitle("Body Size ") +
  theme(legend.position="right", legend.direction="vertical")+
  guides(fill=F, color=F) 
```

## Climate and Wing Length

Load Data
```{r}
climate <- read.csv(file='../data/microclimate/clean/2016TrialsAdultCleaned.csv', stringsAsFactors = F)

climate$Day <- as.Date(climate$Day, format="%Y-%m-%d")
climate$Site_ID <- as.factor(climate$Site_ID)
climate$Tray_ID <- as.factor(climate$Tray_ID)

getClimate <- function(indMosq, climateDF=climate, season){
  #' This is a function to apply over the rows of the octWing and augWing data frames. Must have climate data loaded
  
  #' @param indMosq row of the dataframe for each individual mosquito
  #' @param climateDF the dataframe containing climate data every 10 minutes
  #' @param season, either "summer" or "fall"
  #' @returns formatted data with climate and winglength for the individual mosquito
  
  
  #get date range
  startDate <- ifelse(season=="summer", "2016-08-01", "2016-09-26")
  startDate <- as.Date(startDate, format="%Y-%m-%d")
  endDate <- indMosq$Date
  
  #subset temperature data
  try(climSubset <- climateDF %>%
    filter(Tray_ID==as.character(indMosq$Tray_ID)) %>%
    filter(Day>startDate & Day<endDate),
    silent=T)
  
  #now take mean temperature
  tempMean <- climSubset %>%
    summarise(Tmean=mean(Temp, na.rm=T))
  
  # if (nrow(climSubset)<1000){
  #   climSubset <- climate %>%
  #     filter(Site_ID==indMosq$site) %>%
  #     filter(Day>startDate & Day<endDate)
  # }
  
  #merge this all together
  #mosqFormat <- cbind(indMosq[,c('block','class','site','Tray_ID','Exp_Day', 'mm')], Tmean=tempMean$Tmean)
  
  return(tempMean$Tmean)
}

augWing$Temp <- NA
for(i in 1:nrow(augWing)){
  indMosq <- augWing[i,]
  augWing$Temp[i] <- getClimate(indMosq, season="summer")
}

octWing$Temp <- NA
for(i in 1:nrow(octWing)){
  indMosq <- octWing[i,]
  octWing$Temp[i] <- getClimate(indMosq, season="fall")
}

#combine
fallWing <- octWing %>%
  dplyr::select(block, class, site, Tray_ID, Exp_Day, mm, Temp)

summerWing <- augWing %>%
    dplyr::select(block, class, site, Tray_ID, Exp_Day, mm, Temp)

allWing <- rbind(summerWing, fallWing)
allWing$block<- as.factor(allWing$block)

#drop outlier in S2 (wing size =1.56 mm)
allWing <- allWing %>%
  filter(mm>1.6)
```


```{r}
ggplot(data=allWing, aes(x=Temp, y=mm)) +
  geom_point(aes(col=factor(class), shape=factor(block))) + 
  geom_smooth(method="auto")+
  theme_fivethirtyeight() +
  theme(panel.background = element_rect(fill = "transparent", colour = NA), 
        plot.background = element_rect(fill = "transparent", colour = NA),
        legend.key = element_blank())+
  theme(legend.background = element_rect(fill = "transparent", colour = NA)) +
    theme(axis.title = element_text(), axis.title.x = element_text()) +
  ylab("Uninfected Wing Size")+
  xlab("Mean Temperature (C)") +
  scale_color_manual(values=c(colRural, "gray40", colU), name="Class", labels=c("Rural", "Suburban", "Urban"))+
  #scale_alpha_discrete(range=c(1,0.3), name="Season", labels=c("Fall", "Summer"), guide=guide_legend(reverse=T))+
  scale_shape_discrete(name="Season",labels=c("Fall", "Summer"), guide=guide_legend(reverse=T)) +
  theme(legend.position="right", legend.direction="vertical")+
  guides(shape=F) 
```

# Growth Rate

```{r}
#calculate means per Exp_Day and Tray
growthWing <- allWing %>%
  dplyr::group_by(block,class,site,Tray_ID, Exp_Day) %>%
  dplyr::summarise(meanWing=mean(mm, na.rm=T)) %>%
  dplyr::ungroup()

growthWing$class <- tolower(growthWing$class)

#merge with emergence data
growthDF <- merge(allEmerg, growthWing, by.x=c("block", "Class", "Site_Code", "Tray_Code", "Exp_Day"), by.y=c("block", "class", "site", "Tray_ID", "Exp_Day"), all.x=T) 

trayMeans <- allWing %>%
  dplyr::group_by(block, Tray_ID) %>%
  dplyr::summarise(meanWing=mean(mm, na.rm=T))

#fill in missing with mean of tray during that block
for (i in 1:nrow(growthDF)){
  if (is.na(growthDF$meanWing[i])){
    temp <- trayMeans$meanWing[trayMeans$block==growthDF$block[i] & trayMeans$Tray_ID==growthDF$Tray_Code[i]]
    if (length(temp)==0) next
    growthDF$meanWing[i] <- temp
  } else next
}

#some trays had no mosquitoes emerge that weren't infected, so we take the site level mean for them

siteMeans <- allWing %>%
  dplyr::group_by(site, block) %>%
  dplyr::summarise(meanWing=mean(mm, na.rm=T))

for (i in 1:nrow(growthDF)){
  if (is.na(growthDF$meanWing[i])){
    temp <- siteMeans$meanWing[siteMeans$block==growthDF$block[i] & siteMeans$site==growthDF$Site_Code[i]]
    if (length(temp)==0) next
    growthDF$meanWing[i] <- temp
  } else next
}

```

```{r}
growthDF$Fwx <- -121.240 + (78.02 * growthDF$meanWing)
growthDF$AxFwx <- growthDF$Num_Emerge*growthDF$Fwx
growthDF$xAxFwx <- growthDF$Exp_Day*growthDF$AxFwx

#get sum per day

growthDF2 <- growthDF %>%
  dplyr::group_by(block, Tray_Code, Class, Site_Code) %>%
  dplyr::summarise(xAxFwx = sum(xAxFwx), AxFwx=sum(AxFwx))

growthDF2 <- growthDF2 %>%
  mutate(r=(log((1/50)*AxFwx))/(14+(xAxFwx/AxFwx)))

levels(growthDF2$block) <- c("Summer", "Fall")
levels(growthDF2$Class) <- c("Rural", "Suburban", "Urban")
```

Plot
```{r}
#png(file="figures/forMS/growthRateFig.png", width = 6, height=4, units="in", res=500, family="sans")
ggplot(data=growthDF2, aes(x=Class, y=r))+
  geom_boxplot(aes(fill=Class), alpha=0.5)+
  geom_jitter(aes(color=Class), width=0.1, height=0)+
  facet_wrap(~block)+
  theme_fivethirtyeight() +
  theme(panel.background = element_rect(fill = "transparent", colour = NA), 
        plot.background = element_rect(fill = "transparent", colour = NA),
        legend.key = element_blank())+
  theme(legend.background = element_rect(fill = "transparent", colour = NA)) +
    theme(axis.title = element_text(), axis.title.x = element_text()) +
  ylab("Per Capita Growth Rate (r')")+
  xlab("Land Class") +  
  scale_color_manual(values=c(colR, "gray30", colU)) +
  scale_fill_manual(values=c(colR, colS, colU)) +
  theme(legend.position="right", legend.direction="vertical")+
  guides(fill=F, color=F) 
#dev.off()
```

Statistics:
```{r}
growthMix <- lmer(r~ block + Class + (1|Site_Code), data=growthDF2)
summary(growthMix)
anova(growthMix) #only block is significant
plot(growthMix)
confint(growthMix)
```

# Vectorial Capacity

Arrange temperature data by hour for rate summation
```{r}
climS <- climate[climate$Day >="2016-08-01" & climate$Day <="2016-09-03",]
climS$Block <- "summer"

climF <- climate[climate$Day >="2016-09-26" & climate$Day <="2016-11-08",]
climF$Block <- "fall"

#bind back together
climAll <- rbind(climS, climF)
climAll$Date <- as.POSIXct(climAll$Date)
climAll$Hour <- lubridate::hour(climAll$Date)

parameters <- dplyr::select(climAll, Block, Class, Site_ID, Tray_ID, Temp, Day, Hour)
```

Calculated set at 27C constant (x=27):

Briere:  y ~ a * x * (x - t0) * (tmax - x)^(1/2)
Quad: y ~ a * (x-t0) * (x-Tmax)

```{r}
parameters <- unique(dplyr::select(climAll, Block, Class, Site_ID))
#a -- bite rate
parameters$a <- ((1.93/10000)*27*(27-10.25)*((38.32-27)^0.5))
#adjust negatives
parameters$a[parameters$a<0] <- 0
parameters$a[is.na(parameters$a)] <- 0

# PDR - parasite development rate
parameters$PDR <- ((1.09/10000)*27*(27-10.39)*((43.05-27)^0.5))
#adjust negatives
parameters$PDR[parameters$PDR<0] <- 0
parameters$PDR[is.na(parameters$PDR)] <- 0

# lf - mosquito lifespan
parameters$lf <- -1.43*(27- 13.41)*(27-31.51)
#adjust for zeros
parameters$lf[parameters$lf<0] <- 0
parameters$lf[is.na(parameters$lf)] <- 0

# we will then calculate the mean and se for these parameters per site and hour, then sum them up for a daily value
library(dplyr)

paramRate <- parameters

paramRate <- dplyr::rename(paramRate, Site_Code=Site_ID)
paramRate <- dplyr::rename(paramRate, block=Block)

## u - daily probability of mosquito mortality
paramRate$mu <- 1/paramRate$lf

paramRate$block <- tolower(paramRate$block)
paramRate$Class <- tolower(paramRate$Class)

paramRateOld <- paramRate
```

Combine with field measured fecundity, survival, development rate and vector competence:

```{r}
#EFD: from wing length
EFD  <- allWing %>%
  group_by(block, Class=class, Site_Code=site) %>%
  dplyr::summarise(wingL=mean(mm, na.rm=T)) %>%
  ungroup() %>%
  mutate(fecundity=-121.240 + (78.02*wingL))
EFD$block <- tolower(EFD$block)
EFD$Class <- tolower(EFD$Class)

#pEA: larval survival
pEA <- survSumm %>%
  group_by(block, Class, Site_Code) %>%
  dplyr::summarise(pEA=mean((percSurv/50), na.rm=T))
pEA$block <- tolower(pEA$block)
pEA$Class <- tolower(pEA$Class)

#MDR: emergence rate (day^-1)
MDR <- emergExpF %>%
  group_by(block, Class, Site_Code) %>%
  dplyr::summarise(emerDay=mean(Exp_Day, na.rm=T)) %>%
  mutate(MDR=1/emerDay)
MDR$block <- tolower(MDR$block)
MDR$Class <- tolower(MDR$Class)

#bc
bc <- seasonInfSite %>%
  select(block, Class=class, Site_Code=site, bc=Saliva_mean)
bc$block <- tolower(bc$block)
bc$Class <- tolower(bc$Class)
```

Merge traits together
```{r}
paramRate <- full_join(paramRate, EFD, by=c("block", "Class", "Site_Code"))
paramRate <- full_join(paramRate, pEA, by=c("block", "Class", "Site_Code"))
paramRate <- full_join(paramRate, MDR, by=c("block", "Class", "Site_Code"))
paramRate <- full_join(paramRate, bc, by=c("block", "Class", "Site_Code"))
```

Calculate EFD
```{r}
paramRate$EFD <- paramRate$fecundity*paramRate$a
```

Calcuate VC w/o carry-over effects
```{r}
paramNoCOE <- dplyr::select(climAll, Block, Class, Site_ID, Tray_ID, Temp, Day, Hour)

paramNoCOE$fecundity2 <- ((4.88/100)*27*(27-8.02)*((35.65-27)^0.5))/24
paramNoCOE$fecundity2[paramNoCOE$fecundity2<0] <- 0
paramNoCOE$fecundity2[is.na(paramNoCOE$fecundity2)] <- 0

paramNoCOE$bc2 <- (((7.35/10000)*27*(27-15.84)*((36.40-27)^0.5)) * ((4.39/10000)*27*(27-3.62)*((36.82-27)^0.5)))/24
paramNoCOE$bc2[paramNoCOE$bc2<0] <- 0
paramNoCOE$bc2[is.na(paramNoCOE$bc2)] <- 0

#take mean of each hour and sum up
library(dplyr)
paramRateNoCOE <- paramNoCOE %>%
  group_by(Block, Class, Site_ID, Hour) %>%
  select(-Temp, -Day) %>%
  summarise_each(funs(mean(.,na.rm=T))) %>%
  ungroup() %>%
  group_by(Block, Class, Site_ID) %>%
  summarise_each(funs(sum)) %>%
  ungroup()

paramRateNoCOE <- dplyr::rename(paramRateNoCOE, Site_Code=Site_ID)
paramRateNoCOE <- dplyr::rename(paramRateNoCOE, block=Block)

paramRateNoCOE$Class <- tolower(paramRateNoCOE$Class)
```

Merge w/ and w/o carry-over effects together
```{r}
paramAll <- full_join(paramRate, paramRateNoCOE, by=c("block", "Class", "Site_Code"))

#calculate EFD
#calculate EFD 
paramAll$EFD2 <- paramAll$fecundity2*paramAll$a
```

Calculate VC from traits (VC2 is w/o carry-over)
```{r}
paramAll <- mutate(paramAll, 
                    VC=(((a^2)*bc*(exp(-mu/PDR))*EFD*pEA*(MDR^2))/((mu^3)))^(1/2))

paramAll <- mutate(paramAll,
                       VC2=(((a^2)*bc2*(exp(-mu/PDR))*EFD2*pEA*(MDR^2))/((mu^3)))^(1/2))
```

Add Temperature Back in For Comparison Plots
```{r}
siteTemps <- select(seasonInfSite, block, site, class, Temp_mean_mean)
levels(paramAll$Class) <- c("Rural", "Suburban", "Urban")
VCplot <- merge(paramAll, siteTemps, by.x=c("block", "Site_Code"), by.y=c("block", "site"))
VCplot <- VCplot %>%
  dplyr::select(block, Class, Site_Code, VC, VC2, Temp_mean_mean) %>%
  mutate(VCdiff=(VC-VC2)/VC2*100)
VCplot$block <- factor(VCplot$block, levels=c("summer", "fall"))
```

Comparison Line Plot
```{r}
#png(file="figures/forMS/VCxTemp.png", width = 4, height=4, units="in", res=500, family="sans")
VCLinePlot <- ggplot(data=VCplot, aes(x=Temp_mean_mean))+
  geom_point(aes(y=VC, color="WithCOE")) +
  geom_point(aes(y=VC2, color="WithoutCOE"))+
  geom_smooth(aes(y=VC, color="WithCOE"), method="lm")+
  geom_smooth(aes(y=VC2, color="WithoutCOE"),  method="lm")+
  theme_fivethirtyeight()+
  theme(axis.title = element_text()) +
  theme(panel.background = element_rect(fill = "transparent", colour = NA), 
        plot.background = element_rect(fill = "transparent", colour = NA),
        legend.key = element_blank())+
  theme(legend.background = element_rect(fill = "transparent", colour = NA))+
  theme(legend.position="bottom", legend.direction="horizontal")+
  xlab("Temperature(C)")+
  ylab("Vectorial Capacity")+
  scale_colour_manual(name="VC Calcuation", values=c(WithCOE="dodgerblue", WithoutCOE="maroon"), labels=c("With Carry-Over", "Without Carry-Over"))
VCLinePlot
#dev.off()
```

```{r}
#get mean and se summary
VCplotsumm <- VCplot %>%
  dplyr::select(-Site_Code) %>%
  group_by(block, Class) %>%
  dplyr::summarise_each(funs(mean=mean(., na.rm=T), se=sd(.)/sqrt(n())))
```

Panel Plot with Comparison Line and Bar Graphs
```{r}
png(file="figures/forMS/VCCompareVertical.png", width = 3, height=5, units="in", res=500, family="sans")

#first plot: without COE
plotNoCOE <- ggplot(data=VCplotsumm, aes(x=factor(Class), group=block)) +
  geom_bar(stat='identity',aes(y=VC2_mean, fill=Class, alpha=block), color="black", position=position_dodge(0.9), width=0.7) + 
  geom_errorbar(aes(ymin=VC2_mean-VC2_se, ymax=VC2_mean+VC2_se), position=position_dodge(0.9), width=0.4)+
  theme_base()+
  ylab("Vectorial Capacity")+
  #ggtitle("Effect of COEs on vectorial capacity") +
  xlab("")+
  scale_x_discrete(labels=c("Rural", "Suburban", "Urban"))+
  scale_fill_manual(values=c(colR, colS, colU), labels=c("Rural", "Suburban", "Urban")) +
  scale_alpha_discrete(range=c(1,0.3), name="Season", guide=guide_legend())+
  labs(fill=c("Rural", "Suburban", "Urban"))+
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x=element_blank(), axis.title.y=element_text(size=10))

#second plot: with COE
plotCOE <- ggplot(data=VCplotsumm, aes(x=factor(Class), group=block)) +
  geom_bar(stat='identity',aes(y=VC_mean, fill=Class, alpha=block), color="black", position=position_dodge(0.9), width=0.7) + 
  geom_errorbar(aes(ymin=VC_mean-VC_se, ymax=VC_mean+VC_se), position=position_dodge(0.9), width=0.4)+
  theme_base()+
  ylab("Vectorial Capacity")+
  ylim(0,60)+
  #ggtitle("Effect of COEs on vectorial capacity") +
  xlab("Class")+
  scale_x_discrete(labels=c("Rural", "Suburban", "Urban"))+
  scale_fill_manual(values=c(colR, colS, colU), labels=c("Rural", "Suburban", "Urban")) +
  scale_alpha_discrete(range=c(1,0.3), name="Season", guide=guide_legend())+
  labs(fill=c("Rural", "Suburban", "Urban"))+
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x=element_blank(), axis.title.y=element_text(size=10))

#third plot: difference
plotDiff <- ggplot(data=VCplotsumm, aes(x=factor(Class), group=block)) +
  geom_bar(stat='identity',aes(y=VCdiff_mean, fill=Class, alpha=block), color="black", position=position_dodge(0.9), width=0.7) + 
  #geom_errorbar(aes(ymin=meandiff-sediff, ymax=meandiff+sediff), position=position_dodge(0.9), width=0.4)+
  theme_base()+
  ylab("Percent Change\nin VC")+
  #ggtitle("Effect of COEs on vectorial capacity") +
  xlab("Class")+
  scale_x_discrete(labels=c("Rural", "Suburban", "Urban"))+
  scale_fill_manual(values=c(colR, colS, colU), labels=c("Rural", "Suburban", "Urban")) +
  scale_alpha_discrete(range=c(1,0.3), name="Season", guide=guide_legend(), labels=c("Summer", "Fall"))+
  labs(fill=c("Class","Rural", "Suburban", "Urban"), alpha=c("Season", "Summer", "Fall"))+
  geom_hline(aes(yintercept=0))+
  theme(legend.position="bottom", axis.title.y=element_text(size=10), axis.title.x = element_text(size=12), legend.text=element_text(size=6), legend.title = element_text(size=8), legend.key.size = unit(0.1, "in"), legend.direction = "horizontal", legend.box = "vertical")
  

#legend
plotLegend <- get_legend(plotDiff)

# get rid of legend in third
plotDiff <- plotDiff + theme(legend.position = "none")

#cowplot them together
pgrid <- plot_grid(plotNoCOE, plotCOE, plotDiff, plotLegend, nrow=4, rel_heights = c(1,1,1.2,0.4))

pgrid
# dev.off()
```

VC Stats:
```{r}
VCclass <- lmer(VC~Class*block+ (1|Site_Code), data=VCplot)
plot(VCclass)
summary(VCclass)
car::Anova(VCclass)

#make long for comparison of with and without carry-over
VCtempMix <- glm(VC~Temp_mean_mean, data=VCplot)
```


